Refinements for open automata

### 2022-07-14 2021-Refinement/paper/refinementarticle.tex
Removed solved TODOs and added an explanation to prevent a likely (IMO) confusion

\begin{document}
%
\title{Refinements for open automata}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Rabéa Ameur-Boulifa\inst{1}\orcidID{0000-0002-2471-8012} \and
Quentin Corradi\inst{2}\orcidID{} \and
Ludovic Henrio \inst{2}\orcidID{0000-0001-7137-3523} \and
Eric Madelaine \inst{3}\orcidID{0000-0002-5552-5993}}
%
\authorrunning{Rab\'ea Ameur-Boulifa et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{LTCI, T\'el\'ecom Paris, Institut Polytechnique de Paris, France
\email{first.last@telecom-paris.fr}\\
\and
Universit\'e Lyon, EnsL, UCBL, CNRS, Inria,  LIP France\\
\email{first.last@ens-lyon.fr}
 \and
INRIA Sophia Antipolis M\'edit\'erann\'ee, UCA,  Sophia Antipolis, France
\email{first.last@inria.fr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%\TODO{ The abstract should briefly summarize the contents of the paper in 150--250 words.}

@@

\TODO{Entete : rabea}

\TODO{explication bisim: Ludo}

@@

The notion of refinement  captures the relation between  a specification and an implementation of the same component. Refinement entails that one system can be considered as a more precise version of the specification, featuring all the specified behaviours with more concrete details. 
 It is usually defined as trace inclusion or simulation  \cite{Milner:1980,Kouchnarenko:2007}; this ensures that all behaviours of the specification must belong to the behaviours of the implementation. 
 This definition, which is based on systems whose behaviour is fully defined, is not well-suited for open systems,  as it requires to reason also about unspecified behaviours.

@@

This section presents our notations and the principles of automata. Except for minor changes in the notations, compared to previous works~\cite{fhbisim} the only new contribution is the definition of a composition operator for open automata.
%
%Notations will be defined with the operator \(\defnotation\) and names are given with the operator \(\defobject\) as follows:
%\begin{align*}
%	\mathit{notation\_with\_variables} & \defnotation \mathit{notated\_object\_using\_the\_variables} \\
%	\mathit{name} & \defobject \mathit{fully\_defined\_mathematical\_object}
%\end{align*}

@@

We use two operators to access pieces of information of the OA.

@@

\paragraph{Open automaton composition.}

@@

\begin{definition}[Composition of open automata] \label{Def:CompOA}
% TODO: this definition is a bit strange, why is Ac in the hole of Ap but the hole is in Ac and the notation is reversed?
The composition of \(A_p = \OAg[p]\) in the hole \(k \in J_c\) of    \(A_c = \OAg[c]\) is the OA defined as follows:
\[A_c\subst{A_p}{k} ::=  \OA{S_c \times S_p}{\mpar{s_{0c}, s_{0p}}}{V_c \uplus V_p}{\sigma_{0c} \uplus \sigma_{0p}}{J_p \uplus J_c \setminus \mbrc{k}}{T} \] \text{with }
\begin{align*}
T\!=& \mset{\OT{\mpar{s_c, s_p}}{\mpar{s'_c, s'_p}}{\alpha_c}{\beta_j^{j \in J'_p \uplus J'_c}\!}{g_c \wedge g_p \wedge \alpha_p = \beta_k}{\psi_c \uplus \psi_p}}{ \OT{s_c}{s'_c}{\alpha_c}{\beta_j^{j \in J'_c \uplus \mbrc{k} }\!}{g_c}{\psi_c} \in\! T_c, \OTx{p}{}{}{p} \in\! T_p} \\
	& \cup \mset{\OT{\mpar{s_c, s_p}}{\mpar{s'_c, s_p}}{\alpha_c}{\beta_j^{j \in J'_c}}{g_c}{\psi_c}}{\OTx{c}{}{}{c} \in T_c, k \notin J'_c, s_p \in S_p}
\end{align*}
\end{definition}

@@

\begin{figure}[h]
 \centering
   \includegraphics[width=.5\textwidth]{Figures/counter.pdf}\hfill 
   \includegraphics[width=.3\textwidth]{Figures/compose_count_db.pdf}
   \caption{(Left) Open Automaton of \texttt{counter} process.  Its role is to get a target, then count ticks until that target is reached, then emits an action with the elapsed time and restart. (Right) Open Automaton of the composition \texttt{database}[\texttt{counter/cnt}].\label{Fig:counter}}
\end{figure}
In Fig. \ref{Fig:counter} (right) we present the composition \texttt{database}[\texttt{counter/cnt}] of the two open automata \texttt{database} and \texttt{counter}. Each state of the composite consists of a state of \texttt{database} together with a state of \texttt{counter}. The composed automaton  takes over  the same steps as the   \texttt{database} automaton but it also includes new steps,  indicating the change of states for the setting of timer.

\paragraph{Relations between open automata.}
A relation (bisimulation, refinement, etc.) between OA requires to compare their states. To do so we will suppose that the variables of the two OA are disjoint (a renaming of variables may have to be applied before comparing OA states).
\begin{definition}[Relation on OA configurations] Suppose $V_1$ and $V_2$ are disjoint.
A relation on configurations of \(\OAg[1]\) and \(\OAg[2]\) is a function \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
\end{definition}
%\TODO{C'est etrange, la formulation "the variables they refer to", comme il n'y a plus de variables d'etat... +1}
The idea is that two states are related depending on the state of the automaton, i.e. if the variables of the OA verify a certain formula. 
In other words, for each pair of states, a boolean formula that may refer to the  variables of each of the two OA states whether the two states are related or not.
Additionally, we may check that initial states of the automata are related by checking that: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\).

@@

\paragraph{A bisimulation for open automata.}

Bisimulation between OA was defined in~\cite{fhbisim}. Instead of recalling its definition we show below the principles of this bisimulation.
\TODO{Je ne comprends pas cette figure... elle est incomplete ? Ou alors il faut un peu d'explications...}

@@

Similarly to FH-bisimulation~\cite{fhbisim} we are interested  in finding relations between configurations of two open automata that contain variables and holes. However here we want to build a refinement relation that  also guarantees that no deadlock is introduced when refining the automaton (at least not among the scenarios originally planned in the specification).
The refinement relation should be conditioned by the internal states of the automata. More precisely, a refinement relation characterizes when two states are related, and this  characterisation is expressed as a predicate on the variables of the two automata.
Thus, a refinement relation is a relation on OA configurations.

@@

%The idea is that 
We rely on a classical notion of simulation inspired from~\cite{fhbisim}. The idea is that two configurations related by a relation; if one state can do a transition, then the other can do a transition too. But to limit the creation of deadlocks when simulating an automaton, we first define when a relation is preserved along identical transitions before characterising a first simple refinement relation  to compare automata with identical holes.

@@

Let us first explain the reasons why simulation can introduce allow undesired behaviours to emerge and how we want to limit them. 
The automaton that simulates a specification features any behaviour that contains the behaviour of the specification, or more precisely, for each state that simulates a state of the specification, the simulating automaton must at least feature the behaviour of the specification.
 In this definition, nothing prevents the implementation from adding non-determinism and non-deterministically reaching a state that is unrelated to the specification, this state could even be a deadlock even though we only followed transitions that exist in the specification.
To avoid the appearance of such deadlocks, we will state that, when both the implementation and the specification do the same transition, they stay related.
This ensures a form of continuity in the comparison in the sense that the refined system is constrained to follow the behaviour of the simulated system along a given scenario. 
In terms of traces this amounts to some form of deadlock reduction: for each trace followed by the simulated system if the refined system followed the beginning of the trace, it must be able to keep following the trace. 
Of course, the classical trace view is a simplified view compared to the guards, states and actions from  the holes that exist in OA, but it gives the right intuition.

@@

We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. This leads us to the following definition.
\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
  is preserved along identical transitions if  it satisfies the following:
% TODO: isn't it a2 = a1x instead of a1x = a1
% TODO: ANS:  we have a1 = a2 so whatever, it is the same, we can change if it is more ``natural''
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} \!=\! \OTx{1}{x}{1x}{1x} \!\in\! \fOT{s_1}}^{x \in X}\!,
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\!\fvars{t_{1x}} \uplus \fvars{t_2} \uplus \fvars{t_1} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}}\\
\end{multline*}
\end{definition}

\TODO{[LUDO]
We define here simulation between labelled transition systems to Hole-equal simulation  between  open automata that have exactly the same holes. }
\begin{definition}[Hole-equal simulation]
Consider two OAs \(\OAg[1]\) and \(\OAg[2]\) with  \(J_1 = J_2\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $S_1$ to $S_2$ if the following conditions hold: 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} t_1 = \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{t_{2x} = \OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[10pt]
		\quad \mpar{\forall x \in X, J'_{2x}= J'_1} \nwedge\\[2pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\
\hspace{5em} \left( R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x}} \beta_{1j} = \beta_{2xj} \nwedge\\[12pt]
			 g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}\right)
	\end{array} 
%\\
%	\wedge \mpar{V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}}
\end{multline*}
\item[(3)] $R$ is  is preserved along identical transitions of \(A_1\) and \(A_2\).
\end{definition}

Note in this definition, instead of matching an instantiated transition (with all variables assigned) of the first automata to another instantiated transition of the second, it matches an open transition $t_1$ to a family of covering open transitions $t_{2x}^{x\in X}$. %Note also the second part of the conjunction of condition (2) expresses the deadlock reduction between automata (as introduced in Definition \ref{def:dpwd}).   
The first and second conditions coincide with the natural way to prove inductively that an automaton simulates another by starting with the initial state. The third condition ensures that the refinement relation prevents the introduction of deadlocks.

Intuitively, this means that for every pair of related
states $(s_1,s_2)$  of the two automata, and for every  transition of the first automaton from $s_1$, there is a set of matching transitions  of the second automaton  from $s_2$ such that the produced action match, the actions of the same holes and the successors are related after variable update. Our definition captures a simple kind of sub-classing of open automata with the same holes. It is stronger than a strict simulation since it matches a transition with a family of transitions. 
With such a relation we are able to check the refinement between two open automata with the same level of abstraction but specified differently, for example, by duplicating states, removing transitions,  reinforcing  guards, modifying variables. 
\begin{example} 
To illustrate the refinement simulation of open automata, consider a variation on the \texttt{database} automaton of Fig. \ref{Fig:RefineOA}. We suppose that on more concrete level, in addition to querying and updating the data base,  a backup copy of it may also be made.
\end{example}
\begin{figure}[!tb]
 \centering
   \includegraphics[width=.9\textwidth]{Figures/databaseRefine.pdf}
   \caption{The refined Open Automaton of the automaton of Fig.\ref{Fig:SpecOA}.  Two other actions,  in addition to  \texttt{qry} and \texttt{upd},  \texttt{back} representing a request to make a backup copy and    \texttt{cpy} is used to reflect the actual operation of copying. \label{Fig:RefineOA}}
\end{figure}

The refinement relation  also specifies that the refined process must follow the behaviour of the specification, not only at every compatible state but also along any scenario of the specification.
We will show in Section~\ref{sec:prop} that this refinement relation has good properties in terms of transitivity, compositionality, and reflexivity.

@@

@@

In the following definition we add a parameter $H$ which is the set of holes tracked by the refinement relation and adapt the definition by ignoring actions of the holes that are not in $H$.
\begin{definition}[Open automata refinement]\label{Def:OA-Refinement}
For two open automata\\ \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a refinement of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation on configurations $R: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \nwedge\\[1pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\\hspace{5em} \mpar{R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj}   \nwedge\\
			 g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] $R$ is preserved along identical transitions of \(A_1\) and \(A_2\).

@@

\TODO{[LUDO] relire et virer le rouge}
This definition leads to a notion of simulation as refinement the same as the one used in LOTOS \cite{Brinksma:87} and presented under the name extension refinement.
The extension consists in adding new functionalities by preserving existing ones. Indeed, in case there is no hole and no guard, i.e., in case the definitions of open automata coincide with those of automata, the relation allows new behaviour to be added in terms of new traces, but doesn't allow new deadlocks to
be added, i.e., for traces of the specification (abstract automaton) the implementation (refined automaton) must not add
deadlocks.

@@

\begin{lemma}[Tracked holes] By construction, if an automaton is the refinement of another one, it is also a refinement by tracking less holes.
\[\wrel{A_1}{A_2}{H} \land H'\subseteq H \implies \wrel{A_1}{A_2}{H'}\]
\end{lemma}

\section{Properties of our Refinement Relations}\label{sec:prop}

>>>>

\begin{document}
%
\title{Refinements for open automata}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Rabéa\inst{1}\orcidID{1111-2222-3333-4444} \and
Quentin\inst{1,2}\orcidID{1111-2222-3333-4444} \and
Ludo\inst{2}\orcidID{0000-1111-2222-3333} \and
Eric\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{Rab\'ea Ameur-Boulifa et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{blabla
\email{lncs@fff}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
blibli\\
\email{\{abc,def\}@fff}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%\TODO{ The abstract should briefly summarize the contents of the paper in 150--250 words.}

@@

@@

The notion of refinement  captures the relation between  a specification and an implementation of the same component. Refinement entails that one system can be considered as a more precise version of the specification, featuring all the specified behaviours with more concrete details. 
It is usually defined as trace inclusion or simulation  \cite{Milner:1980,Kouchnarenko:2007}; this ensures that all behaviours of the specification must belong to the behaviours of the implementation.
This definition, which is based on systems whose behaviour is fully defined, is not well-suited for open systems,  as it requires to reason also about unspecified behaviours.
% TODO: This version of simulation/trace inclusion doesn't fit our needs for another reason,
% in our case all behaviours of the implementation do actually belong to the behaviours of the specification with our refinement (it seems counter-intuitive so it may be worth explaining):
% Our specifications says "you must do that the way you choose to do it" (more precisely "the way the automaton in this hole chooses to do it as long as the guard is satisfied"), the implementation says "I do that this specific way"
% so the behaviour of the implementation is a behaviour allowed by the specification, in particular it is included in the behaviours (of/allowed by) the specification

@@

This section presents our notations and the principles of automata. Except for minor changes in the notations, compared to previous works~\cite{pnets} the only new contribution is the definition of a composition operator for open automata.
%
%Notations will be defined with the operator \(\defnotation\) and names are given with the operator \(\defobject\) as follows:
%\begin{align*}
%	\mathit{notation\_with\_variables} & \defnotation \mathit{notated\_object\_using\_the\_variables} \\
%	\mathit{name} & \defobject \mathit{fully\_defined\_mathematical\_object}
%\end{align*}

@@

@@

\paragraph{Open automaton composition}

@@

\begin{definition}[Composition of open automata] \label{Def:CompOA}
The composition of \(A_p = \OAg[p]\) in the hole \(k \in J_c\) of    \(A_c = \OAg[c]\) is the OA defined as follows:
\[A_c\subst{A_p}{k} ::=  \OA{S_c \times S_p}{\mpar{s_{0c}, s_{0p}}}{V_c \uplus V_p}{\sigma_{0c} \uplus \sigma_{0p}}{J_p \uplus J_c \setminus \mbrc{k}}{T} \] \text{with }

\begin{align*}
T=& \mset{\OT{\mpar{s_c, s_p}}{\mpar{s'_c, s'_p}}{\alpha_c}{\beta_j^{j \in J'_p \uplus J'_c}}{g_c \wedge g_p \wedge \alpha_p = \beta_k}{\psi_c \uplus \psi_p}}{ \OT{s_c}{s'_c}{\alpha_c}{\beta_j^{j \in J'_c \uplus \mbrc{k} }}{g_c}{\psi_c} \in T_c, \OTx{p}{}{}{p} \in T_p} \\
	& \cup \mset{\OT{\mpar{s_c, s_p}}{\mpar{s'_c, s_p}}{\alpha_c}{\beta_j^{j \in J'_c}}{g_c}{\psi_c}}{\OTx{c}{}{}{c} \in T_c, k \notin J'_c, s_p \in S_p}
\end{align*}
\end{definition}

@@

\begin{figure}[h]
 \centering
   \includegraphics[width=.5\textwidth]{Figures/counter.pdf}\hfill 
   \includegraphics[width=.3\textwidth]{Figures/compose_count_db.pdf}
   \caption{(left) Open Automaton of \texttt{counter} process.  Its role is to get a target, then count ticks until that target is reached, then emits an action with the elapsed time and restart. (right) Open Automaton of the composition \texttt{database}[\texttt{counter/cnt}].\label{Fig:counter}}
\end{figure}
In Fig. \ref{Fig:counter} (right) we present the composition \texttt{database}[\texttt{counter/cnt}] of the two open automata \texttt{database} and \texttt{counter}. Each state of the composite consists of a state of \texttt{database} together with a state of \texttt{counter}. The composed automaton  takes over  the same steps as the   \texttt{database} automaton but it also includes new steps,  indicating the change of states for the setting of timer.

\paragraph{Relations between open automata}
A relation (bisimulation, refinement, etc.) between OA requires to compare their states. To do so we will suppose that the variables of the two OA are disjoint (a renaming of variables may have to be applied before comparing OA states).
\begin{definition}[Relation on OA states] Suppose $V_1$ and $V_2$ are disjoint.
A relation on states of \(\OAg[1]\) and \(\OAg[2]\) is a function \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
\end{definition}
\TODO{C'est etrange, la formulation "the variables they refer to", comme il n'y a plus de variables d'etat... +1}
The idea is that two states are related if the variables they refer to verify a certain formula. Additionally, we may check that initial states of the automata are related by checking that: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\).

@@

\paragraph{A bisimulation for open automata}

Bisimulation between OA was defined in~\cite{fh-bisim}. Instead of recalling its definition we show below the principles of this bisimulation.
\TODO{Je ne comprends pas cette figure... elle est incomplete ? Ou alors il faut un peu d'explications...}

@@

Similarly to FH-bisimulation~\cite{fhbisim} we are interested  in finding relations between states of two open automata that contain variables and holes. However here we want to build a refinement relation that  also guarantees that no deadlock is introduced when refining the automaton (at least not among the scenarios originally planned in the specification).
The refinement relation should be conditioned by the internal states of the automata. More precisely, a refinement relation characterizes when two states are related, and this  characterisation is expressed as a predicate on the variables of the two automata.
Thus, a refinement relation is a relation on OA states.

@@

%The idea is that 
We rely on a classical notion of simulation inspired from~\cite{fh-bisim}. The idea is that two states related by a given relation and states that if one state can do a transition, then the other can do a transition too. But to limit the creation of deadlocks when simulating an automaton, we first define when a relation is preserved along identical transitions before characterising a first simple refinement relation  to compare automata with identical holes.

@@

Let us first explain the reasons why simulation can introduce allow undesired bwhaviours to emerge and how we want to limit them. 
The automaton that simulates a specification features any behaviour that contains the behaviour of the specification, or more precisely, for each state that simulates a state of the specification, the simulating automaton must at least feature the behaviour of the specification.
 In this definition, nothing prevents the implementation from adding non-determinism and non-deterministically reaching a state that is unrelated to the specification, this state could even be a deadlock even though we only followed transitions that exist in the specification.
To avoid the appearance of such deadlocks, we will state that, when both the implementation and the specification do the same transition, they stay related.
This ensures a form of continuity in the comparison in the sense that the refined system is constrained to follow the behaviour of the simulated system along a given scenario. 
In terms of traces this amounts to some form of deadlock reduction: for each trace followed by the simulated system if the refined system followed the beginning of the trace, it must be able to keep following the trace. 
Of course, the classical trace view is a simplified view compared to the guards, states and actions from  the holes that exist in OA, but it gives the right intuition.

@@

We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. This leads us to the following definition.
\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
  is preserved along identical transitions if  it satisfies the following:
% TODO: isn't it a2 = a1x instead of a1x = a1
% TODO: ANS:  we have a1 = a2 so whatever, it is the same, we can change if it is more ``natural''
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} = \OTx{1}{x}{1x}{1x} \in \fOT{s_1}}^{x \in X},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_2} \uplus \fvars{t_1} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}}\\
\end{multline*}
\end{definition}

\medskip

We define here simulation between labelled transition systems to Hole-equal simulation  between  open automata that have exactly the same holes.
\begin{definition}[Hole-equal simulation]
Consider two OAs \(\OAg[1]\) and \(\OAg[2]\) with  \(J_1 = J_2\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $S_1$ to $S_2$ if the following conditions hold: 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} t_1 = \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{t_{2x} = \OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[10pt]
		\quad \mpar{\forall x \in X, J'_{2x}= J'_1} \nwedge\\[2pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\
\hspace{5em} \left( R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x}} \beta_{1j} = \beta_{2xj} \nwedge\\[12pt]
			 g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}\right)
	\end{array} 
%\\
%	\wedge \mpar{V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}}
\end{multline*}
\item[(3)] $R$ is  is preserved along identical transitions of \(A_1\) and \(A_2\).
\end{definition}

Note in this definition, instead of matching an instantiated transition (with all variables assigned) of the first automata to another instantiated transition of the second, it matches an open transition $t_1$ to a family of covering open transitions $t_{2x}^{x\in X}$. %Note also the second part of the conjunction of condition (2) expresses the deadlock reduction between automata (as introduced in Definition \ref{def:dpwd}).   
\TODO{The first and second conditions coincide with the natural way to prove inductively that an automaton simulates another by starting with the initial state. The third condition ensures that the refinement relation prevents the introduction of deadlocks.}

Intuitively, this means that for every pair of related
states $(s_1,s_2)$  of the two automata, and for every  transition of the first automaton from $s_1$, there is a set of matching transitions  of the second automaton  from $s_2$ such that the produced action match, the actions of the same holes and the successors are related after variable update. Our definition captures a simple kind of sub-classing of open automata with the same holes. It is stronger than a strict simulation since it matches a transition with a family of transitions. 
With such a relation we are able to check the refinement between two open automata with the same level of abstraction but specified differently, for example, by duplicating states, removing transitions,  reinforcing  guards, modifying variables.

To illustrate the refinement simulation of open automata, consider a variation on the \texttt{database} automaton of Fig. \ref{Fig:RefineOA}. We suppose that on more concrete level, in addition to querying and updating the data base,  a backup copy of it may also be made.

\begin{figure}[h]
 \centering
   \includegraphics[width=.9\textwidth]{Figures/databaseRefine.pdf}
   \caption{The refined Open Automaton of the automaton of Fig.\ref{Fig:SpecOA}.  Two other actions,  in addition to  \texttt{qry} and \texttt{upd},  \texttt{back} representing a request to make a backup copy and    \texttt{cpy} is used to reflect the actual operation of copying. \label{Fig:RefineOA}}
\end{figure}

\TODO{The refinement relation  also specifies that the refined process must follow the behaviour of the specification, not only at every compatible state but also along any trace.}

 We will show in Section~\ref{sec:prop} that this refinement relation has good properties in terms of transitivity, compositionality, reflexivity, etc.

@@

\TODO{add an example}

@@

In the following definition we add a parameter $H$ which is the set of holes tracked by the refinement relation and adapt the definition by ignoring actions of the holes that are not in $H$.
\begin{definition}[Open automata refinement]\label{Def:OA-Refinement}
For two open automata\\ \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a refinement of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation $R: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \nwedge\\[1pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\\hspace{5em} \mpar{R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj}   \nwedge\\
			 g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] $R$ is preserved along identical transitions of \(A_1\) and \(A_2\).

@@

This definition leads to a notion of simulation as refinement the same as the one used in LOTOS \cite{Brinksma:87} and presented under the name extension refinement.
The extension consists in adding new functionalities by preserving existing ones. Indeed, in case there is no hole and no guard, i.e., in case the definitions of open automata coincide with those of automata, the relation allows new behaviour to be added in terms of new traces, but doesn't allow new deadlocks to
be added, i.e., for traces of the specification (abstract automaton) the implementation (refined automaton) must not add
deadlocks.

@@

\begin{lemma}[Tracked holes] By construction, if an automaton is the refinement of another one, it is also a refinement by tracking less holes.
\[\wrel{A_1}{A_2}{H} \land H'\subseteq H \implies \wrel{A_1}{A_2}{H'}\]
\end{lemma}
\TODO{example}

\section{Properties of our Refinement Relations}\label{sec:prop}
\TODO{1.5 pages}
We now  state the properties of our refinement relations, proofs of the properties can be found in the appendices. We express these properties in terms of open automata refinement because hole-equal simulation is a particular case of Definition~\ref{Def:OA-Refinement} when  $J_1=J_2=H$, and thus most of the properties shown here are easy to adapt to hole-equal simulation.
